Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression\thanksref{T1}

### 2020-11-02 revision_2/response/Drafts of response to referees/response.tex

\item As a methodological paper, what’s the guideline/recommendation given by the authors? After reading this paper, I don’t know how to apply the suggested framework
on a simple regression model. Apparently, the choice of the likelihood function (for the good data) $f(y|\theta)$ and the choice of the statistic T are related.
To make things more complicated, each has multiple choices. For example, f could be normal, student-t or other heavy-tailed distribution; T could be something named Huber or Tukey although the authors are not even bothered to provide any mathematical expression.

Instead of just reporting numerical performance, the authors should provide some guidelines/recommendation on how to apply their framework on linear regression models, e.g., what are the default choices for f and T ? Is there any- way to select which f or T to use based on cross-validation or other empirical methods?

\response{Thanks for pointing out this need. We have taken the following steps:

\begin{enumerate}
\item Point out that M-Estimators have been our default choice in the very first illustrative example:
 \response{Details of these estimators can be found in many places, including \citep{huber2009}. We return to the two M-estimators throughout this paper as we have found them to offer good default choices for practitioners dealing with outliers. A short review of these estimators is provided in the Supplementary Material.}
 \item Included a short review of M-Estimators in the Supplement
 \item Included a `Practical Considerations for Using the Restricted Likelihood' to the Supplement. 
\end{enumerate}
}

@@

\response{At the end of section 4 we have included a paragraph:

`
\response{Finally, it is clear the estimators themselves must be computed for every iteration of the Markov Chain. We have found this burden to be marginal in relation with respect to computing the needed Jacobian. In the simulations and real data analyses presented below, we will see that the additional burden is often of added value over traditional robust regression when substantial prior information is available that is not swamped by current data.}
`
Additionally we have included a paragraph in the introduction discussing advantages and disadvantages.
}

\item The authors compare KL diver-gence, but 1) few real applications care about KL divergence not mentioning the predictive density f(y|M,yobs) is not available in closed-form; 2) since f(y|M,yobs) is not available close-form, the KL divergence is approximated and it’s not clear to me how accurate the approximation is.
The authors should compare prediction and estimation accuracy, common met- rics used for regression models.

\response{We have changed to a closed form version of KL between two normals to address estimation concern you bring up: $ 
\log(\hat\sigma^{2}) - \log(\sigma^{2}) + \frac{1}{\hat\sigma^{2}}(\sigma^{2} + (\theta - \hat{\theta_{i}})^{2}
$
While we wouldn't be able to compute this in a real application it does combines estimation error for both the $\theta_{i}$ and $\sigma$ and is a decent metric to compare the methods under the simulation setup. }

@@

`
\response{These M-estimators satisfy C3 and C4 since they are optimizers of (almost surely) continuous and differentiable objective functions. Constraints C5-C8 are often satisfied by location and scale estimators but should be checked on a case by case basis.}
`
}

\item  Bottom of p14, ”Sample z? from a distribution with known density on S.” Any distribution? For example, can it be a point mass or its support has to be equal to S? What’s the sampling distribution for z? used in the simulation studies and real data analysis?

\response{Thanks for noticing this detail. It should have full support on S since the transformation is 1-1 and onto A. We have now mentioned this. We have also included in discussion of what we used for p throughout based on a previous comment. This comment is after Theorem 4.6.}

\item  Eq (18): change dy to d$\tilde{y}$?
\response{Thanks for noticing this typo - based on your other comment on KL divergence, this formula was taken out and now is a closed form version. }

### 2020-11-02 revision_2/Work on the revision early files/paperV3.tex
clean up and add abc analysis script

\begin{frontmatter}
\title{Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression\thanksref{T1}}
\runtitle{Bayesian Restricted Likelihood Methods}
\thankstext{T1}{This research has been supported by Nationwide Insurance Company and by the NSF under grant numbers DMS-1007682 and DMS-1209194.  The views in this paper are not necessarily those of Nationwide Insurance or the NSF.}

@@

The remainder of the paper is as follows: Section~\ref{restrictedlikelihood} introduces the Bayesian restricted likelihood, provides context with previous work, and demonstrates some advantages of the methods on simple examples. Section~\ref{BayesLinMod} details an MCMC algorithm to apply the method to Bayesian linear models. This computational strategy is a major contribution to the work, providing an approach to apply the method on realistic examples. Many of the the technical proofs are in the Appendix \ref{sec:appendix} with \texttt{R} code available from the authors. Sections \ref{simData} and \ref{RealData} illustrate the method with simulated data and a real insurance industry data set containing many outliers with a novel twist on model evaluation. A discussion (Section~\ref{Conclusions}) provides some final commentary on the new method. \response{ An R package \texttt{brlm} to implement our methods is available at \texttt{github.com/jrlewi/brlm}. Additionally all data and code for the examples in this paper are available at \texttt{https://github.com/jrlewi/brlm\char`_paper/revision\char`_2}.}

@@

The second example involves deliberate censoring of small and large observations. This is
sometimes done as a precursor to the analysis of reaction time experiments  \citep[e.g.,][]{ratcliff1993} where very small and large reaction times are physiologically implausible;  explained by either anticipation or lack of attention of the subject.  
With lower and upper censoring times at $t_1$ and $t_2$, the post-censoring sampling distribution is of mixed form, with masses $F(t_1|\bth)$ at $t_1$ and $1-F(t_2|\bth)$ at $t_2$,
and density $f(y | \bth)$ for $y \in (t_1, t_2)$.  We adjust the original data $y_i$,
producing $c(y_i)$ by defining $c(y_i)= t_1$ if $y_i \leq t_1$, $c(y_i)=t_2$ 
if $y_i \geq t_2$, and $c(y_i)=y_i$ otherwise.  
The adjusted update is performed with $L(\bth |c(\by))$.  
Letting $g(t_1|\bth) = F(t_1|\bth)$,
$g(t_2 | \bth) = 1 - F(t_2|\bth)$, and $g(y|\bth)=f(y|\bth)$ for
$y \in (t_1, t_2)$, we may rewrite the full 
likelihood as the product of two pieces
\begin{eqnarray}
\label{Censoring} 
L(\bth  | \by) =  \left( \prod_{i=1}^n g(c(y_i)  | \bth) \right) \left( \prod_{i=1}^n f(y_i|\bth,c(y_i)). \right) ,  
\end{eqnarray}

$\prod_{i=1}^n f(y_i|\bth,c(y_i))$ is the likelihood of the data conditioned on parameters and the summary statistic $c(\cdot)$ and recovers the piece of the full likelihood not in $ \prod_{i=1}^n g(c(y_i)  | \bth)$. Only the first part is retained in the analysis. Several more examples are detailed in \cite{lewis2014}.

@@

Given the one-to-one and onto mapping $h: \mathbb{S} \rightarrow \mathcal{A}$, the general proposal strategy is summarized as follows:
\begin{enumerate}
\item Sample $\bz^*$ from a distribution with known density \response{whose support is the entirety} $\mathbb{S}$.
\item Set $\by = h(\bz^*)$ and calculate the Jacobian of this transformation in two steps.
\begin{enumerate}
\item Scale from $\mathbb{S}$ to the set $\Pi(\mathcal{A}):= \{\bz\in \mathbb{R}^n |\ \exists\ \by\in \mathcal{A}\ s.t.\ \bz=Q \by \}$ with $Q = I - XX^{\top}$. \footnote{We have used condition \ref{fullRank} to assume  without loss of generality  that the columns of $X$ form an orthonormal basis for $\mc{C}(X)$ (i.e., $X^\top X=I$).} $\Pi(\mathcal{A})$ is the projection of $\mathcal{A}$ onto $\mathcal{C}^{\perp}(X)$ and, by condition \ref{regIn}, every element of this set has $s(X, \bz) = s(X, \by_{obs})$. Specifically, set $\bz=\frac{s(X,\boldsymbol{y}_{obs})}{s(X, \boldsymbol{z}^{*})}\bz^{*}$. There are two pieces of this Jacobian: one for the scaling and one for the mapping of the sphere onto $\Pi(\mathcal{A})$. The latter piece is given in equation \eqref{cosine}.
\item Shift  from $\Pi(\mathcal{A})$ to $\mathcal{A}$: $\by=\bz+X\left(\bb(X, \by_{obs})-\bb(X, \bz)\right)$. This shift is along the column space of $X$ to the unique element in $\mathcal{A}$. The Jacobian of this transformation is given by equation \eqref{eq:volume}.
\end{enumerate}
\end{enumerate}

@@

\response{To asses the predictive capability, we seek a metric that takes into account both the estimation of $\theta_{i}$ and $\sigma^{2}$. In this model, estimation of $\theta_{i}$ and prediction are analogous. To this end we consider the loss:
\begin{equation}
\label{sim1loss}
L((\hat\theta_{i}, \hat\sigma_{i}), (\theta_{i},\sigma)) =
\log(\hat\sigma_{i}^{2}) - \log(\sigma^{2}) + \frac{1}{\hat\sigma_{i}^{2}}(\sigma^{2} + (\theta_{i} - \hat{\theta_{i}})^{2})
\end{equation}
Note, this loss is proportional to the Kullback-Leibler divergence between two normals with the base normal being the true distribution of the good portion of the data and the the other normal representing the estimate under the fitting method. We are using fixed point estimates $\hat\theta_{i}$ and $\hat\sigma$ which are posterior means under the Bayesian models and the standard point estimates for the classical methods.
}
%To assess predictive capability, the models are compared using Kullback-Leibler (KL) divergence from the distribution of good data to the posterior predictive distribution. Specifically, for the $i^{th}$ group of the $k^{th}$ simulated data set $\by_{k}$ compute:
%\label{kl}
%KL^{(M)}_{ik} = \int \log \frac{f(\tilde y | \theta_{i}, \sigma^{2})}{f_{i}(\tilde y | M, \by_{k})}  f(\tilde y | \theta_{i}, \sigma^{2}) \ dy
%where $M$ indexes the fitting model and $f(\tilde y | \theta_{i}, \sigma^{2}) = N(\tilde y | \theta_{i}, \sigma^{2})$, the normal density function with (known) mean $\theta_{i}$ and variance $\sigma^{2}$, evaluated at $\tilde y$. For the Bayesian models, ${f_{i}(\tilde y | M, \by_{k})} = \int f(\tilde y |\theta_{i}, \sigma_{i}^{2}) \pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})d\theta_{i}d\sigma^{2}_{i}$ where $\pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})$ is the posterior for the $i^{th}$ group model parameters under model $M$ for the $k^{th}$ data set. $M$ denotes either the full normal theory model \eqref{fullsim2} or one of the two restricted likelihood versions, along with specified $a_{s}$ and $c$. For the classical robust fits, we set $f_{i}(\tilde y | M, \by_{k}) = N(\tilde y|\hat\theta_{i}, \hat\sigma^{2}_{i})$ as a groupwise plug-in estimator for the predictive distribution. The classical fits are computed separately for each group with no consideration of the hierarchical structure between the groups. The overall mean $\overline{KL}^{(M)}_{{\cdot}{\cdot}} = \frac{1}{90K} \sum_{k = 1}^{K} \sum_{i=1}^{90} KL^{(M)}_{ik}$ is used to compare the models, where smaller means correspond to better fits. Sampling variation is summarized with the standard error between the $K=30$ replicates in the simulation:  $SE(\overline{KL}^{(M)}_{{\cdot} k}) = \sqrt{\frac{1}{K(K-1)}\sum_{k = 1}^{K} (\overline{KL}^{(M)}_{{\cdot} k} - \overline{KL}^{(M)}_{{\cdot}{\cdot}})^{2}}$ where $\overline{KL}^{(M)}_{{\cdot} k} = \frac{1}{90}\sum_{i = 1}^{90} KL^{(M)}_{ik}$.
Figure \ref{kl_sim} displays the average of loss \ref{sim1loss} grouped by pairs of $a_{s}$ and $c$ with error bars plus/minus on standard error within the group.  The values of $a_{s}$ and $c$, do not affect the classical robust linear models. The average loss for the normal theory models ranges from $1.44$ to $1.66$ and are left out of the figure. For $c = 0.5$ and $c = 1$ the results favor the restricted likelihood methods as seen by the lower values of loss. The choice of $c = 2$ corresponds to a particularly poor prior distribution.  The prior has substantial mass above $\sigma^{2} = 4$, with prior means for $\sigma^2$ from $8.9$ to $32$ as $a_s$ varies.  Additionally, the tuning parameters chosen for the location and scale estimators result in an upward bias in the estimate of $\sigma^{2}$. This bias depends on $m$ and $p$. For example, for $m = 9$ and $p = .1$, Huber's version converges to roughly $4.8$ as $n$ grows.  The bias is greater for more severe levels of contamination.  The alignment of biases in prior distribution and in likelihood from the summary statistic (when applied to the contaminated data) inflates the estimate of scale.  Not surprisingly, a poor prior distribution whose weakness matches the weakness in the likelihood results in poorer inference.  In this case, poorer than the classical estimators.

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{mse_sim2_facet_scale.png}
\caption{Average Loss plus/minus one standard error for each value of $a_{s}$ and $c$. Smaller values represent better fits. The panels correspond to $c = 0.5$ (left), $c=1$ (middle), and $c=2$ (right), with the values of $a_{s}$ on the horizontal axis. The average loss for the normal theory model ranges from $1.44$ to $1.66$ and is left out of the figure.}
\label{kl_sim}
\end{figure}

It is also interesting to consider the effects of factors $n$, $p$, and $m$.  We present the results for a single prior ($a_s = 5$ and $c = 1$).  For each simulation $k$, the main effect average loss is found for each factor $n$, $p$, and $m$. Figure \ref{kl_mnp} displays the average of these main effects over the $K = 30$ simulations along  with error bars plus/minus one standard error. For each group $n$, $p$, and $m$, the Bayesian restricted likelihood versions have better average loss than do the classical methods. As expected, the average loss gets larger as the contamination gets more severe (larger $m$ or larger $p$) and tends to get smaller as the sample size $n$ grows.  The advantage of the Bayesian method is greater for smaller sample sizes.

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{MSE_sim2_mnp.png}
\caption{Average loss plus/minus one standard error grouped by the factors $m$ (left), $n$ (middle), and $p$ (right). These results are for the single prior with $a_s = 5$ and $c = 1$.}
\label{kl_mnp}
\end{figure}

This simulation shows the potential of the restricted likelihood and conveys some cautions.  Specifically, the choice of summary statistics, along with corresponding tuning parameters is important. For the tuning parameters, we applied the default choice of $95\%$ efficiency at the normal.  Under the simulation model here, this choice results in bias in the scale estimation which affects the performance of the method. These choices must be made when using both the classical and Bayesian methods. The Bayesian approach encourages use of a hierarchical model structure and allows one to incorporate prior information in the analysis.  These features can improve predictive performance substantially.  If poorly handled, they can, of course, harm performance.

@@

In this simulation the data are generated from the following mechanism: $y = \bbeta^{\top}x + \epsilon$
with $\bbeta = (\beta_{1}, \beta_{2},\beta_{3})^{\top}$ and the error $\epsilon \sim N(0,\sigma^{2})$ with probability $0.8$ and $\epsilon \sim \text{Half-Normal}(0,25\sigma^{2})$ with probability $0.2$ (i.e., there is a relatively large amount of one-sided outlier contamination). The components of $x = (x_{1}, x_{2}, x_{3})$ are correlated with $x_1 \sim N(0,1)$ and $x_{j} = x_{1} + \eta_{j}$ with $\eta_{j} \sim N(0, 4)$ for $j = 2,3$. This results in a theoretical correlation of $1/sqrt(5) \approx 0.44$ between $x_{1}$ and both $x_{j}, j = 2, 3$. The model used for fitting contains an additional 27 covariates, some of which are also correlated with $x_{1}$, $x_{2}$, and $x_{3}$. Specifically the fitting model is $y = \beta^{\top} x + \beta^{*\top} x^{*} + \epsilon$ where $x^{*}$ and $\beta^{*}$ are 27 dimensional vectors of extra covariates and slope parameters. Of these 27 covariates, 21 are generated independently from standard normal distributions. Of the remaining 6, two each are generated by adding standard normal noise to $x_{1}$, $x_{2}$, and $x_{3}$. This represents a common situation where several covariates with various levels of correlation amongst them are available for fitting, but only a few govern the data generating mechanism.

>>>>

\begin{frontmatter}
\title{Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression\thanksref{T1}}
\runtitle{Bayesian Restricted Likelihood Methods}
\thankstext{T1}{This research has been supported by Nationwide Insurance Company and by the NSF under grant numbers DMS-1007682, DMS-1209194, DMS-15-13566, and DMS-16-13110.  The views in this paper are not necessarily those of Nationwide Insurance or the NSF.}

@@

The remainder of the paper is as follows: Section~\ref{restrictedlikelihood} introduces the Bayesian restricted likelihood, provides context with previous work, and demonstrates some advantages of the methods on simple examples. Section~\ref{BayesLinMod} details an MCMC algorithm to apply the method to Bayesian linear models. This computational strategy is a major contribution to the work, providing an approach to apply the method on realistic examples. Many of the the technical proofs are in the Appendix \ref{sec:appendix} with \texttt{R} code available from the authors. Sections \ref{simData} and \ref{RealData} illustrate the method with simulated data and a real insurance industry data set containing many outliers with a novel twist on model evaluation. A discussion (Section~\ref{Conclusions}) provides some final commentary on the new method. \response{ An R package \texttt{brlm} to implement our methods is available at \texttt{github.com/jrlewi/brlm}. Additionally all data and code for the examples in this paper are available at \texttt{https://github.com/jrlewi/brlm\char`_paper/}.}

@@

The second example involves deliberate censoring of small and large observations. This is
sometimes done as a precursor to the analysis of reaction time experiments  \citep[e.g.,][]{ratcliff1993} where very small and large reaction times are physiologically implausible;  explained by either anticipation or lack of attention of the subject.  
With lower and upper censoring times at $t_1$ and $t_2$, the post-censoring sampling distribution is of mixed form, with masses $F(t_1|\bth)$ at $t_1$ and $1-F(t_2|\bth)$ at $t_2$,
and density $f(y | \bth)$ for $y \in (t_1, t_2)$.  We adjust the original data $y_i$,
producing $c(y_i)$ by defining $c(y_i)= t_1$ if $y_i \leq t_1$, $c(y_i)=t_2$ 
if $y_i \geq t_2$, and $c(y_i)=y_i$ otherwise.  
The adjusted update is performed with $L(\bth |c(\by))$.  
Letting $g(t_1|\bth) = F(t_1|\bth)$,
$g(t_2 | \bth) = 1 - F(t_2|\bth)$, and $g(y|\bth)=f(y|\bth)$ for
$y \in (t_1, t_2)$, we may rewrite the full 
likelihood as the product of two pieces
\begin{eqnarray}
\label{Censoring} 
L(\bth  | \by) =  \left( \prod_{i=1}^n g(c(y_i)  | \bth) \right) \left( \prod_{i=1}^n f(y_i|\bth,c(y_i)). \right) ,  
\end{eqnarray}
$\prod_{i=1}^n f(y_i|\bth,c(y_i))$ is the likelihood of the data conditioned on parameters and the summary statistic $c(\cdot)$ and recovers the piece of the full likelihood not in $ \prod_{i=1}^n g(c(y_i)  | \bth)$. Only the first part is retained in the analysis. Several more examples are detailed in \cite{lewis2014}.

@@

Given the one-to-one and onto mapping $h: \mathbb{S} \rightarrow \mathcal{A}$, the general proposal strategy is summarized as follows:
\begin{enumerate}
\item Sample $\bz^*$ from a distribution with known density \response{whose support is the entirety of} $\mathbb{S}$.
\item Set $\by = h(\bz^*)$ and calculate the Jacobian of this transformation in two steps.
\begin{enumerate}
\item Scale from $\mathbb{S}$ to the set $\Pi(\mathcal{A}):= \{\bz\in \mathbb{R}^n |\ \exists\ \by\in \mathcal{A}\ s.t.\ \bz=Q \by \}$ with $Q = I - XX^{\top}$. \footnote{We have used condition \ref{fullRank} to assume  without loss of generality  that the columns of $X$ form an orthonormal basis for $\mc{C}(X)$ (i.e., $X^\top X=I$).} $\Pi(\mathcal{A})$ is the projection of $\mathcal{A}$ onto $\mathcal{C}^{\perp}(X)$ and, by condition \ref{regIn}, every element of this set has $s(X, \bz) = s(X, \by_{obs})$. Specifically, set $\bz=\frac{s(X,\boldsymbol{y}_{obs})}{s(X, \boldsymbol{z}^{*})}\bz^{*}$. There are two pieces of this Jacobian: one for the scaling and one for the mapping of the sphere onto $\Pi(\mathcal{A})$. The latter piece is given in equation \eqref{cosine}.
\item Shift  from $\Pi(\mathcal{A})$ to $\mathcal{A}$: $\by=\bz+X\left(\bb(X, \by_{obs})-\bb(X, \bz)\right)$. This shift is along the column space of $X$ to the unique element in $\mathcal{A}$. The Jacobian of this transformation is given by equation \eqref{eq:volume}.
\end{enumerate}
\end{enumerate}

@@

\response{The performance of the methods can be evaluated in many ways.  For these simulations, we know the true data generating mechanism, and this allows us to make direct comparisons between the fitted model and the truth.  The Bayesian methods provide a full predictive distribution for the response, given group while the classical methods provide only point estimates of parameters.  Our comparisons have focused on two main summaries.  One summary, not presented here, is the average Kullback-Liebler divergence from the good portion of the true distribution of $Y$ given group to the predictive distribution (Bayes) or distribution with point estimates plugged in (classical).  For the Bayesian models, the divergence does not have a closed form and must be evaluated numerically.  The second summary, preferred by a referee, is the mean squared error (MSE), averaged across groups.  Results are presented in Figures \ref{mse_sim} and \ref{mse_mnp}.  Formally, with the superscript $M$ indicating the method and the additional subscript $k$ indexing the data set,
%\begin{equation}
\begin{align}
\label{MSE}
\mbox{MSE}^{(M)}_{ik} & = (\hat\theta^{(M)}_{ik} - \theta_{ik})^2,  \\ 
\mbox{MSE}^{(M)} & = \frac{1}{90K} \sum_{k = 1}^{K} \sum_{i=1}^{90} \mbox{MSE}^{(M)}_{ik}
\end{align}
%\end{equation}

The Bayesian restricted likelihood methods show superior performance under both summaries, but especially for MSE.  Figure \ref{mse_sim} displays the MSE grouped by pairs $a_{s}$ and $c$ with error bars plus/minus one standard error within the group.  The values of $a_{s}$ and $c$, do not affect the classical robust linear models. The average MSEs for the normal theory models ranges from $0.24$ to $0.25$ and are left out of the figure.  The results uniformly favor the Bayesian restricted likelihood methods, as seen by substantially lower values of MSE.  For both classical and restricted likelihood methods, Tukey's $\psi$ function leads to better performance than does Huber's $\psi$ function.  }
%For $c = 0.5$ and $c = 1$ the results favor the restricted likelihood methods as seen by the lower values of loss. The choice of $c = 2$ corresponds to a particularly poor prior distribution.  The prior has substantial mass above $\sigma^{2} = 4$, with prior means for $\sigma^2$ from $8.9$ to $32$ as $a_s$ varies.  Additionally, the tuning parameters chosen for the location and scale estimators result in an upward bias in the estimate of $\sigma^{2}$. This bias depends on $m$ and $p$. For example, for $m = 9$ and $p = .1$, Huber's version converges to roughly $4.8$ as $n$ grows.  The bias is greater for more severe levels of contamination.  The alignment of biases in prior distribution and in likelihood from the summary statistic (when applied to the contaminated data) inflates the estimate of scale.  Not surprisingly, a poor prior distribution whose weakness matches the weakness in the likelihood results in poorer inference.  In this case, poorer than the classical estimators. .  }

%\response{To asses the predictive capability, we seek a metric that takes into account both the estimation of $\theta_{i}$ and $\sigma^{2}$. In this model, estimation of $\theta_{i}$ and prediction are analogous. To this end we consider the loss:
%\begin{equation}
%\label{sim1loss}
%L((\hat\theta_{i}, \hat\sigma_{i}), (\theta_{i},\sigma)) =
%\log(\hat\sigma_{i}^{2}) - \log(\sigma^{2}) + \frac{1}{\hat\sigma_{i}^{2}}(\sigma^{2} + (\theta_{i} - \hat{\theta_{i}})^{2})
%\end{equation}
%Note, this loss is proportional to the Kullback-Leibler divergence between two normals with the base normal being the true distribution of the good portion of the data and the the other normal representing the estimate under the fitting method. We are using fixed point estimates $\hat\theta_{i}$ and $\hat\sigma$ which are posterior means under the Bayesian models and the standard point estimates for the classical methods.
%}
%To assess predictive capability, the models are compared using Kullback-Leibler (KL) divergence from the distribution of good data to the posterior predictive distribution. Specifically, for the $i^{th}$ group of the $k^{th}$ simulated data set $\by_{k}$ compute:
%\begin{equation}
%\label{kl}
%KL^{(M)}_{ik} = \int \log \frac{f(\tilde y | \theta_{i}, \sigma^{2})}{f_{i}(\tilde y | M, \by_{k})}  f(\tilde y | \theta_{i}, \sigma^{2}) \ dy
%\end{equation}
%where $M$ indexes the fitting model and $f(\tilde y | \theta_{i}, \sigma^{2}) = N(\tilde y | \theta_{i}, \sigma^{2})$, the normal density function with (known) mean $\theta_{i}$ and variance $\sigma^{2}$, evaluated at $\tilde y$. For the Bayesian models, ${f_{i}(\tilde y | M, \by_{k})} = \int f(\tilde y |\theta_{i}, \sigma_{i}^{2}) \pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})d\theta_{i}d\sigma^{2}_{i}$ where $\pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})$ is the posterior for the $i^{th}$ group model parameters under model $M$ for the $k^{th}$ data set. $M$ denotes either the full normal theory model \eqref{fullsim2} or one of the two restricted likelihood versions, along with specified $a_{s}$ and $c$. For the classical robust fits, we set $f_{i}(\tilde y | M, \by_{k}) = N(\tilde y|\hat\theta_{i}, \hat\sigma^{2}_{i})$ as a groupwise plug-in estimator for the predictive distribution. The classical fits are computed separately for each group with no consideration of the hierarchical structure between the groups. The overall mean $\overline{KL}^{(M)}_{{\cdot}{\cdot}} = \frac{1}{90K} \sum_{k = 1}^{K} \sum_{i=1}^{90} KL^{(M)}_{ik}$ is used to compare the models, where smaller means correspond to better fits. Sampling variation is summarized with the standard error between the $K=30$ replicates in the simulation:  $SE(\overline{KL}^{(M)}_{{\cdot} k}) = \sqrt{\frac{1}{K(K-1)}\sum_{k = 1}^{K} (\overline{KL}^{(M)}_{{\cdot} k} - \overline{KL}^{(M)}_{{\cdot}{\cdot}})^{2}}$ where $\overline{KL}^{(M)}_{{\cdot} k} = \frac{1}{90}\sum_{i = 1}^{90} KL^{(M)}_{ik}$.

%Figure \ref{kl_sim} displays the average of loss \ref{sim1loss} grouped by pairs of $a_{s}$ and $c$ with error bars plus/minus on standard error within the group.  The values of $a_{s}$ and $c$, do not affect the classical robust linear models. The average loss for the normal theory models ranges from $1.44$ to $1.66$ and are left out of the figure. For $c = 0.5$ and $c = 1$ the results favor the restricted likelihood methods as seen by the lower values of loss. The choice of $c = 2$ corresponds to a particularly poor prior distribution.  The prior has substantial mass above $\sigma^{2} = 4$, with prior means for $\sigma^2$ from $8.9$ to $32$ as $a_s$ varies.  Additionally, the tuning parameters chosen for the location and scale estimators result in an upward bias in the estimate of $\sigma^{2}$. This bias depends on $m$ and $p$. For example, for $m = 9$ and $p = .1$, Huber's version converges to roughly $4.8$ as $n$ grows.  The bias is greater for more severe levels of contamination.  The alignment of biases in prior distribution and in likelihood from the summary statistic (when applied to the contaminated data) inflates the estimate of scale.  Not surprisingly, a poor prior distribution whose weakness matches the weakness in the likelihood results in poorer inference.  In this case, poorer than the classical estimators.

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{mse_sim2_facet_scale.png}
\caption{Average MSE plus/minus one standard error for each value of $a_{s}$ and $c$. Smaller values represent better fits. The panels correspond to $c = 0.5$ (left), $c=1$ (middle), and $c=2$ (right), with the values of $a_{s}$ on the horizontal axis. The average MSE for the normal theory model ranges from $0.24$ to $0.25$ and is left out of the figure.}
\label{mse_sim}
\end{figure}

\response{It is also interesting to consider the effects of factors $n$, $p$, and $m$.  We present the results for a single prior ($a_s = 5$ and $c = 1$).  For each simulation $k$, the main effect average MSE is found for each factor $n$, $p$, and $m$. Figure \ref{mse_mnp} displays the average MSE over the $K = 30$ simulations along with error bars plus/minus one standard error. For each group $n$, $p$, and $m$, the Bayesian restricted likelihood versions have better average loss than do the classical methods. As expected, the average MSE gets larger as the contamination gets more severe (larger $m$ or larger $p$) and tends to get smaller as the sample size $n$ grows.  The advantage of the Bayesian methods is greater for smaller sample sizes and for less severe contamination.}

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{MSE_sim2_mnp.png}
\caption{Average MSE plus/minus one standard error grouped by the factors $m$ (left), $n$ (middle), and $p$ (right). These results are for the single prior with $a_s = 5$ and $c = 1$.}
\label{mse_mnp}
\end{figure}

\response{This simulation shows the potential of the restricted likelihood methods to dramatically improve estimation.  The simulation also conveys some cautions that are apparent from consideration of KL divergence but not MSE.  Specifically, the choice of summary statistics, along with corresponding tuning parameters is important. For the tuning parameters, we applied the default choice of $95\%$ efficiency at the normal.  Under the simulation model here, this choice results in bias in the scale estimation which affects the performance of the method. These choices must be made when using both the classical and Bayesian methods. The Bayesian approach encourages use of a hierarchical model structure and allows one to incorporate prior information in the analysis.  These features can improve predictive performance substantially.  If poorly handled, they can, of course, harm performance. }

@@

In this simulation the data are generated from the following mechanism: $y = \bbeta^{\top}x + \epsilon$
with $\bbeta = (\beta_{1}, \beta_{2},\beta_{3})^{\top}$ and the error $\epsilon \sim N(0,\sigma^{2})$ with probability $0.8$ and $\epsilon \sim \text{Half-Normal}(0,25\sigma^{2})$ with probability $0.2$ (i.e., there is a relatively large amount of one-sided outlier contamination). The components of $x = (x_{1}, x_{2}, x_{3})$ are correlated with $x_1 \sim N(0,1)$ and $x_{j} = x_{1} + \eta_{j}$ with $\eta_{j} \sim N(0, 4)$ for $j = 2,3$. This results in a theoretical correlation of $1/\sqrt{5} \approx 0.44$ between $x_{1}$ and both $x_{j}, j = 2, 3$. The model used for fitting contains an additional 27 covariates, some of which are also correlated with $x_{1}$, $x_{2}$, and $x_{3}$. Specifically the fitting model is $y = \beta^{\top} x + \beta^{*\top} x^{*} + \epsilon$ where $x^{*}$ and $\beta^{*}$ are 27 dimensional vectors of extra covariates and slope parameters. Of these 27 covariates, 21 are generated independently from standard normal distributions. Of the remaining 6, two each are generated by adding standard normal noise to $x_{1}$, $x_{2}$, and $x_{3}$. This represents a common situation where several covariates with various levels of correlation amongst them are available for fitting, but only a few govern the data generating mechanism.

For the simulation, $K = 30$ data sets (including the additional covariates) of size $n = 500$ are generated from the true model with true values $\bbeta = (1,1,1)^{\top}$ and $\sigma^{2} = 2$. We fit the model including all 30 covariates and consider the following methods for the fit 1) classical robust regression with Tukey's estimator of location and Huber's estimator of scale, 2) the corresponding restricted likelihood version 3) a heavy-tailed Bayesian model with a Student-t likelihood with $\nu = 5$ degrees of freedom.  For the Bayesian models we take $\bbeta_{all}\sim N_{20}(\mb{0}, \sigma^{2}_{\beta}I)$ with $\bbeta_{all} = (\bbeta, \bbeta^{*})^{\top}$ and $\sigma^{2}\sim IG(5,8)$ under the restricted model and  $\sigma^{2}\sim IG(5,\frac{\nu-2}{\nu}8)$ under the Student-t model.  For each data set, we fit the models for $ \sigma_{\beta} = 0.4, 0.6, 0.8,\dots, 1.4$. The acceptance rates for the restricted likelihood MCMC data-augmentation step range from $0.3$ to $0.36$ across all the data sets and values of $\sigma_{\beta}$. To compare performance we first consider the $MSE = (||\bbeta - \hat\bbeta||^{2} + ||\hat\bbeta^{*}||^{2})/30$ for each simulation where $\hat\bbeta$ and $\hat\bbeta^{*}$ are point estimates for the fitted model. For the Bayesian models, we use posterior means. The average MSEs plus/minus one standard error over the simulations for each $\sigma^{2}_{\beta}$ are displayed in Figure \ref{mseSimMany}. The classical fit is labeled `rlm` and is the same for each value of the prior standard deviation $\sigma^{2}_{\beta}$. We see for most values of the the prior standard deviation, the Bayesian models (`restricted' and `t') outperform the classical fit. The correlation amongst the covariates causes a certain level of confounding and the prior shrinkage helps to improve estimation. However, too much shrinkage can be detrimental as demonstrated for $\sigma_{\beta} = 0.4$. While this will help for estimation of $\bbeta^{*} = 0$, the estimation of the active parameters $\bbeta$ can be hindered. The $t$ model seems more sensitive to this effect than the restricted model. The restricted model also has an additional advantage when it comes to prediction of the non-outlying data. To see this, for each simulation we consider the mean negative log-likelihood of the non-outlying data: $MNLL = -\frac{1}{N} \sum \log f(y_{i} |  \hat\bbeta, \hat\bbeta^{*}, \hat\sigma)$ where $f$ is the assumed likelihood and the average is taken over the $N$ non-outlying points $y_{i}$. For the classical and restricted fits, $f$ is the normal likelihood and for the `t' it is the heavy-tailed Student-t likelihood. The average MNLL plus/minus one standard error over the simulations for each $\sigma^{2}_{\beta}$ are displayed in Figure \ref{negllSimMany}. First, the restricted version has a small but consistent improvement over the classical method. Second, it is clear that that the heavy-tailed model suffers when trying to predict the non-outlying data since it assumes the entire data generating mechanism is heavy-tailed.

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{mse_sim_many_p.png}
\caption{Average MSE plus/minus one standard error over the $K = 30$ simulations for each value of the prior standard deviation ($\sigma_{\beta}$) and each of the fitting methods. `Restricted' is our method conditioning on Tukey's estimator of location and Huber's estimator of scale. `rlm' refers to the classical robust linear model fit with the same estimators and `t' is the heavy-tailed Bayesian model with a Student-t likelihood. The `rlm' results are the same for each $\sigma_{\beta}$.}
\label{mseSimMany}
\end{figure}
